---
title: "GloBIAnalysis-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GloBIAnalysis-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GloBIAnalysis)
```

# Package uses

This package is designed to aid naïve users in the analysis of the GloBI plant-bee interaction datasets. Functions include:

-   Summarizing: designed to help the user understand the distribution of the data sets and the biologically-important variables included

-   Visualization: designed to help the user create simple visualizations of the different types of bee-plant interactions that are recorded in the data sets

-   ANOVA: designed to facilitate easy two-way ANOVA analysis of species interactions in relation to the endemicity of the bees or plants involved

# Example uses of functions

## Summarizing the data

I wonder what bees are included in this data?

```{r}
summaryBee()
```

I wonder what plants are included in this data?

```{r}
summaryPlant()
```

I wonder what types of bee-plant interactions are included in this data?

```{r}
summaryInteract()
```

I wonder where in the world this data comes from?

```{r}
summaryGeo()
```

## Visualizing the data

I still don't understand all the different bees included in this data and I want to visualize it.

```{r}
visBeeSpecies()
```

I still don't understand all the different plants included in this data and I want to visualize it.

```{r}
visPlantSpecies()
```

I still don't understand the different types of bee-plant interactions included in this data and I want to visualize it.

```{r}
visInteractType()
```

I want to understand where these interactions are taking place... visually!

```{r}
visInteractMap()
```

## Statistical analysis

I wonder if bee endemicity has a significant relationship with the types of interactions that occur?

```{r}
anovaPop("bee", 0.05)
```

I wonder if plant endemicity has a significant relationship with the types of interactions that occur?

```{r}
anovaPop("plant", 0.01)
```
